Local Spectral Clustering of Density Upper Level Sets

### 2021-01-31 jmlr/revision/archive/local_spectral_density_clustering_jmlr_revision.tex
Final draft to be submitted to JMLR uploaded, in revision/ppr_clustering_neighborhood_graph

%\subsubsection{Sweep cut}
%\label{subsection: sweep_cut}
Next, we define a \emph{$\beta$-sweep cut} of $\pbf_v = (\pbf_v(u))_{u \in V}$,
for a given level $\beta > 0$, as    
\begin{equation}
\label{eqn: sweep_cuts}
S_{\beta,v} := \set{u \in V: \frac{\pbf_v(u)}{\Dbf_{uu}} > \beta}.
\end{equation}
We will use the normalized cut metric to determine which sweep cut $S_{\beta}$
is the best cluster estimate. For a set $S \subseteq V$ with complement $S^c = V
\setminus S$, we define \smash{$\cut(S;G_{n,r}) := \sum_{u \in S, v \in S^c}
\Abf_{uv}$}, and \smash{$\vol(S; G_{n,r}) := \sum_{u \in S} \Dbf_{uu}$}.  We
then define the \emph{normalized cut} of $S$ as
\begin{equation}
\label{eqn: normalized_cut}
\Phi(S; G_{n,r}) := \frac{\cut(S;G_{n,r})}{\min \set{\vol(S; G_{n,r}), \vol(S^c; G_{n,r})}}.
\end{equation}
Having computed sweep cuts $S_{\beta}$ over \smash{$\beta \in (L, U)$} (where
the range $(L,U)$ is user-specified), we output the cluster estimate
\smash{$\Cest = S_{\beta^*}$} with minimum normalized cut.  For concreteness,
the PPR algorithm is summarized in Algorithm~\ref{alg: ppr}.

@@

In general, the study of spectral algorithms on neighborhood graphs has been
focused on establishing asymptotic convergence of eigenvalues and eigenvectors
of certain sample objects to the eigenvalues and eigenfunctions of corresponding
limiting operators. \citet{koltchinskii2000} establish convergence of spectral
projections of the adjacency matrix to a limiting integral operator, with
similar results obtained using simplified proofs in
\citet{rosasco10}. \citet{vonluxburg2008} studies convergence of eigenvectors of
the Laplacian matrix for a neighborhood of fixed radius. \citet{belkin07} and
\citet{garciatrillos18} extend these results to the regime where the radius $r
\to 0$ as $n \to \infty$.

@@

At a high level, for PPR to be successful, the underlying density cluster must
be geometrically well-conditioned.  A basic requirement is that we need to avoid
clusters which contain arbitrarily thin bridges or spikes. As in the work of
\citet{chaudhuri2010}, we consider a thickened version of \smash{$\Cset \in
\Cbb_f(\lambda)$} defined as $\Csig := \set{x \in \Reals^d: \dist(x,\Cset) \leq
\sigma}$, which we call the \emph{$\sigma$-expansion} of $\Cset$. Here
\smash{$\dist(x,\Cset) := \inf_{y \in \Cset} \norm{y - x}$}.  We now list our
conditions on $\Csig$.

\begin{enumerate}[label=(A\arabic*)]
\item
  \label{asmp: bounded_density}
  \emph{Bounded density within cluster:} There exist constants
  $0<\lambda_{\sigma}< \Lambda_{\sigma}<\infty$ such that 
  $$
  \lambda_{\sigma} \leq \inf_{x \in \Csig} f(x) \leq \sup_{x \in \Csig} f(x)
  \leq \Lambda_{\sigma}.
  $$

\item 
  \label{asmp: low_noise_density}
  \emph{Low noise density:} There exists constants $c_0 > 0$ and $\gamma \in
  [0,1]$ such that for any $x \in \Rd$ with $0 < \dist(x, \Csig) \leq \sigma$,     
  $$
  \inf_{x' \in \Csig} f(x') - f(x) \geq  c_0 \cdot \dist(x, \Csig)^{\gamma}.  
  $$
  Roughly, this assumption ensures that the density decays sufficiently quickly
  as we move away from the target cluster $\Csig$, and is a standard assumption
  in the level-set estimation literature (see for instance \citet{singh2009}).

\item
  \label{asmp: embedding}
  \emph{Lipschitz embedding:}
  There exists $g: \Reals^d \to \Reals^d$ with the following properties: 
  \begin{enumerate}
  \item we have $\Csig = g(\mathcal{K})$, for a convex set $\mathcal{K}
    \subseteq \Rd$ with $\mathrm{diam}(\mathcal{K}) = \sup_{x,y \in
      \mathcal{K}}\norm{x - y} =: \diam < \infty$;
  \item $\det(\nabla g (x)) = 1$ for all $x \in \Csig$, where $\nabla g(x)$ is
    the Jacobian of $g$ evaluated at $x;$ and 
  \item for some $L \geq 1$,   
    $$
    \norm{g(x) - g(y)} \leq L \norm{x - y} ~
    \text{for all $x,y \in \mathcal{K}$}. 
    $$
  \end{enumerate}
  Succinctly, we assume that $\Csig$ is the image of a convex set with finite
  diameter under a measure preserving, Lipschitz transformation.

\item
  \label{asmp: bounded_volume}
  \emph{Bounded volume:}
  For a set $\mathcal{S} \subseteq \Reals^d$, define a $\Pbb$-weighted volume of 
  $\mathcal{S}$ to be  
  \begin{equation}
  \label{eqn:volume}
  \vol_{\Pbb,r}(\mathcal{S}) := \int_{\mathcal{S}} \Pbb(B(x,r)) f(x) \, dx.
  \end{equation}
  where $B(x,r)$ is the closed ball of radius $r$ centered at $x$. We assume
  that the radius of the neighborhood graph $0 < r \leq \sigma/(2d)$ is chosen 
  such that  
  $$
  \vol_{\Pbb,r}(\Cset_\sigma) \leq \frac{1}{2} \vol_{\Pbb,r}(\Reals^d). 
  $$
\end{enumerate}

@@

\textcolor{red}{Need to fix this definition: (1) the universal constants are not given in the Theorem statements, and (2) the definition of $\tau_u(\mc{C})$ has changed.}

\begin{definition}
  \label{def:well_conditioned_density_cluster}
  For $\lambda > 0$ and \smash{$\Cset \in \Cbb_f(\lambda)$}, let $\Cset$ satisfy  
  \ref{asmp: bounded_density}--\ref{asmp: bounded_volume} for some $\sigma >
  0$. Then, for universal constants $c_1, c_2, c_3 > 0$ to be specified later
  (in Theorems \ref{thm: conductance_upper_bound} and \ref{thm:
    mixing_time_upper_bound}), define
  \begin{equation}
    \label{eqn: condition_number}
    \Phibf(\Cset) 
    := c_1 r \frac{d}{\sigma} \frac{\lambda}{\lambda_{\sigma}}
    \frac{(\lambda_{\sigma} - c_0 \frac{r^{\gamma}}{\gamma +
        1})}{\lambda_{\sigma}}, \quad
    \taubf(\Cset) := c_2 \frac{\Lambda_{\sigma}^4 d^3 \rho^2
      L^2}{\lambda_{\sigma}^4 r^2}
    \log^2\left(\frac{\Lambda_{\sigma}}{\lambda_{\sigma}^2r}\right) + c_3.
  \end{equation}
  Letting \smash{$\kappa(\Cset) := \Phibf(\Cset) \cdot \taubf(\Cset)$},
  we call $\kappa(\Cset)$ the \emph{condition number} of $\Cset$.  We also call
  the set $\Cset$ a \emph{$\kappa$-well-conditioned} density cluster.  
\end{definition}

The condition number $\kappa(\Cset)$ succinctly captures the role of the various 
geometric parameters.  We note in passing that $\Phibf(\Cset)$ and
$\taubf(\Cset)$ are exactly the upper bounds  on \smash{$\Phi(\Csig[\Xbf];
  G_{n,r})$} and \smash{$\tau_{\infty}(G_{n,r}[\Csig[\Xbf]])$} that we derive in
our analysis later, in Section \ref{sec: analysis}.

@@

\begin{equation}
\begin{gathered}
\label{eqn: initialization}
0 < r \leq \frac{\sigma}{2d}, \quad 
\alpha \in {\textstyle [\frac{1}{10}, \frac{1}{9})} \cdot
\frac{1}{\taubf(\theta)}, \\ 
(L,U) \subseteq {\textstyle(\frac{1}{50},\frac{1}{5})} \cdot 
\frac{1}{2{n \choose 2} \vol_{\Pbb,r}(\Csig)}.  
\end{gathered}
\end{equation}

@@

We now present our first main result: a bound on the volume of the symmetric set
difference between the estimated cluster \smash{$\Cest$} and empirical cluster 
$\Csig[\Xbf]$. In this theorem, and hereafter, we let $c, c_i > 0$ denote
universal constants, and $b,b_i > 0$ denote constants which may depend on  
$\Pbb,\lambda,r,d$ and so on, but \emph{not} on the sample size 
$n$.  % where possible, we keep track of all constants in our proofs.

\textcolor{red}{Need to fix this Theorem.}

\begin{theorem}
  \label{thm: volume_ssd}
  Fix $\lambda,\sigma > 0$, let \smash{$\Cset \in \Cbb_f(\lambda)$} be a   
  $\kappa$-well-conditioned density cluster, and assume Algorithm \ref{alg: ppr}
  is well-initialized with respect to $\Cset$. Then for any 
  $$
  n \geq b_1 (\log n)^{\max\{\frac{3}{d},1\}},
  $$
  there exists a set $\Csig[\Xbf]^g 
  \subseteq \Csig[\Xbf]$ of large volume, \smash{$\vol_{n,r}(\Csig[\Xbf]^g)
    \geq \vol_{n,r}(\Csig[\Xbf])/2$}, such that the following holds: if 
  Algorithm~\ref{alg: ppr} is run with any seed node $v \in \Csig[\Xbf]^g$, then
  the PPR estimated cluster \smash{$\Cest$} satisfies
  \begin{equation}
    \label{eqn: volume_ssd}
    \Delta(\Cest, \Csig[\Xbf]) \leq c \cdot \kappa(\Cset) \cdot
    \vol_{n,r}(\Csig[\Xbf]),  
  \end{equation}
  with probability at least $1 - b_2/n$. 
\end{theorem}

@@

\begin{theorem}
  \label{thm: conductance_upper_bound}
  Fix $\lambda,\sigma>0$, and assume $\Cset \in \Cbb_f(\lambda)$ satisfies    
  Assumptions \ref{asmp: bounded_density}, \ref{asmp: low_noise_density}, and   
  \ref{asmp: bounded_volume}. Then for any $\delta \in (0,1/2)$, it holds that
  \begin{equation}
    \label{eqn: conductance_additive_error_bound}
    \frac{\Phi_{n,r}(\Csig[\Xbf])}{r} \leq (1 + 4\delta) c_1\frac{d}{\sigma}
    \frac{\lambda}{\lambda_{\sigma}} \frac{(\lambda_{\sigma} -
      c_0\frac{r^{\gamma}}{\gamma+1})}{\lambda_{\sigma}},
  \end{equation}
  with probability at least $1 - 3\exp\set{-b n \delta^2}$. 
\end{theorem}

@@

\begin{theorem}
  \label{thm: mixing_time_upper_bound}
  Fix $\lambda, \sigma > 0$, and assume \smash{$\Cset \in \Cbb_f(\lambda)$} 
  satisfies Assumptions \ref{asmp: bounded_density} and \ref{asmp: embedding}.
  Also assume that \smash{$0 < r \leq \sigma/(4d)$}. Then for any \smash{$\delta \in \bigl(0,1/(8 + 16/d)\bigr)$}, and any
  $$
  n \geq b_1 \cdot (1 + 4\delta) \cdot 
  \begin{cases*}
  \log(n)^{3/2},& ~~\textrm{if $d = 2$,} \\
  \log(n),& ~~\textrm{if $d \neq 2$}
  \end{cases*}
  $$
  the mixing time satisfies
  \begin{align} 
    \label{eqn: mixing_time_upper_bound}
    \tau_{\infty}\bigl(G_{n,r}[\Csig[\Xbf]]\bigr) & \leq (1 + 40\delta) c_2 \frac{\Lambda_{\sigma}^4 d^3 
      \rho^2 L^2}{\lambda_{\sigma}^4 r^2} \log\Bigl(c_4^{1/d}\frac{\Lambda_{\sigma}^{2/d}L\rho}{\lambda_{\sigma}^{2/d}\sigma}\Bigr)\log\Bigl(c_5^{1/d}\frac{\Lambda_{\sigma}^{2/d}L\rho}{\lambda_{\sigma}^{2/d}\sigma}\Bigr) + \\
  &~~c_3\biggl((1 + 32\delta)d\log\Bigl(c_4^{1/d}\frac{\Lambda_{\sigma}^{2/d}L\rho}{\lambda_{\sigma}^{2/d}\sigma}\Bigr) + 1\biggr), 
  \end{align}
  with probability at least $1 - b_5/n - 4n\exp\{-\delta^2 b_2 n\}$.
\end{theorem}
The proof of Theorem \ref{thm: mixing_time_upper_bound} relies heavily on
analogous mixing time bounds developed for a continuous-space ``ball walk'' over
convex sets. To the best of our knowledge, our result is the first bound on the
mixing time of random walks over neighborhood graphs that is independent of $n$,
the number of vertices.

@@

%In this supplement, we present proofs and experimental details for the article ``Local Spectral Clustering of Density Upper Level Sets''.
\section{Proof of Theorem \ref{thm: conductance_upper_bound}}
\label{sec: proof_of_theorem_1}

@@

With this notation in place, the goal of Theorem~\ref{thm: conductance_upper_bound} is to show that for a universal constant $c_1 > 0$,
\begin{align*}
\Phi_{n,r}(\Csig[\Xbf]) := \frac{\cut_{n,r}(\Csig[\Xbf])}{\min\{\vol_{n,r}(\Csig[\Xbf]), \vol_{n,r}((\Rd \setminus \Csig)[\Xbf])\}} \leq c_1 \frac{d}{\sigma}
    \frac{\lambda}{\lambda_{\sigma}} \frac{(\lambda_{\sigma} -
      c_0\frac{r^{\gamma}}{\gamma+1})}{\lambda_{\sigma}}
\end{align*}
with probability at least $1 - 3\exp\{-nb\}$.

The proof of this theorem follows essentially from two technical Lemmas. 
Lemma~\ref{lem:ball_bounds_in_probability} relates the terms in the numerator and denominator of $\Phi_{n,r}(\Csig[\Xbf])$ to their expected values. We restate the conclusions of this Lemma: for any $\delta > 0$,
\begin{align}
\frac{\cut_{n,r}(\Csig[\Xbf])}{2{n \choose 2}} \leq (1 + \delta)\cut_{\Pbb,r}(\Csig) &, \quad \frac{\vol_{n,r}(\Csig[\Xbf])}{2{n \choose 2}} \geq (1 - \delta)\vol_{\Pbb,r}(\Csig) \nonumber \\ \frac{\vol_{n,r}((\Rd \setminus \Csig)[\Xbf])}{2{n \choose 2}}& \geq (1 - \delta)\vol_{\Pbb,r}(\Rd \setminus \Csig) \label{eqn: conductance_upper_bound_pf1}
\end{align}
with probability at least 
\begin{align*}
1 - & \exp\set{-n \delta^2 (\cut_{\Pbb,r}(\Csig))^2} - \exp\set{-n \delta^2 (\vol_{\Pbb,r}(\Csig))^2} - \exp\set{-n \delta^2 (\vol_{\Pbb,r}(\Rd\setminus \Csig))^2} \\
& \geq 1 - 3\exp\set{-n \delta^2 (\cut_{\Pbb,r}(\Csig))^2}.
\end{align*} 
We note that as a consequence of~\ref{asmp: bounded_volume} we have that $\vol_{\Pbb,r}(\Rd \setminus \Csig)  \geq \vol_{\Pbb,r}(\Csig)$, so it will suffice to lower bound $\vol_{\Pbb,r}(\Csig)$ (since a lower bound for $\vol_{\Pbb,r}(\Rd \setminus \Csig)$ follows).
The following result provides upper and lower bounds on the expected values $\cut_{\Pbb,r}(\Csig)$ and $\vol_{\Pbb,r}(\Csig)$ respectively:
\begin{lemma}
	\label{lem: expected_density_cut}\label{lem: expected_density_volume}
	Under the setup and conditions of Theorem \ref{thm: conductance_upper_bound}, and for any $0 < r \leq \sigma/(2d)$,
	\begin{align}
	\cut_{\Pbb,r}(\Csig) &\leq \frac{ d \nu_d r^{d+1} \lambda}{\sigma - dr} \left(\lambda_{\sigma} - c_0\frac{r^{\gamma}}{\gamma + 1}\right) \nu(\Csig), \label{eqn:claim_one} \\
	\vol_{\Pbb,r}(\Csig) &\geq \biggl(1 + \frac{dr}{\sigma - dr}\biggr)^{-1} \lambda_{\sigma}^2 \nu_d r^d \nu(\Csig).\label{eqn:claim_two}
	\end{align}
\end{lemma}
Taking Lemma~\ref{lem: expected_density_cut} and \eqref{eqn: conductance_upper_bound_pf1} as given we can now complete the proof of the theorem. We lower bound $\Phi_{n,r}(\Csig[\Xbf])$ as follows:
\begin{equation*}
\Phi_{n,r}(\Csig[\Xbf]) \leq \frac{(1 + \delta)\cut_{\Pbb,r}(\Csig)}{(1 - \delta)\vol_{\Pbb,r}(\Csig)} \leq \frac{(1 + \delta)}{(1 - \delta)(1 - dr/\sigma)^2}\frac{d \nu_d r \lambda \left(\lambda_{\sigma} - c_0\frac{r^{\gamma}}{\gamma + 1}\right)}{\sigma \lambda_{\sigma}^2}.
\end{equation*}
Noting that $(1 - dr/\sigma)^{-2} \leq 1 + 6dr/\sigma \leq 4$ for all $r \leq \sigma/2d$, the theorem is satisfied by choosing $c_1 = 4$ and $b = (\cut_{\Pbb,r}(\Csig))^2$.

>>>>

Next, we define a \emph{$\beta$-sweep cut} of $\pbf_v = (\pbf_v(u))_{u \in V}$,
for a given level $\beta > 0$, as    
\begin{equation}
\label{eqn: sweep_cuts}
S_{\beta,v} := \set{u \in V: \frac{\pbf_v(u)}{\Dbf_{uu}} > \beta}.
\end{equation}
We will use the normalized cut metric to determine which sweep cut $S_{\beta}$
is the best cluster estimate. For a set $S \subseteq V$ with complement $S^c = V
\setminus S$, we define \smash{$\cut(S;G_{n,r}) := \sum_{u \in S, v \in S^c}
\Abf_{uv}$}, and \smash{$\vol(S; G_{n,r}) := \sum_{u \in S} \Dbf_{uu}$}.  We
then define the \emph{normalized cut} of $S$ as
\begin{equation}
\label{eqn: normalized_cut}
\Phi(S; G_{n,r}) := \frac{\cut(S;G_{n,r})}{\min \set{\vol(S; G_{n,r}), \vol(S^c; G_{n,r})}}.
\end{equation}
Having computed sweep cuts $S_{\beta}$ over \smash{$\beta \in (L, U)$} (where
the range $(L,U)$ is user-specified), we output the cluster estimate
\smash{$\Cest = S_{\beta^*}$} with minimum normalized cut.  For concreteness,
the PPR algorithm is summarized in Algorithm~\ref{alg: ppr}.

@@

In general, the study of spectral algorithms on neighborhood graphs has been
focused on establishing asymptotic convergence of eigenvalues and eigenvectors
of certain sample objects to the eigenvalues and eigenfunctions of corresponding
limiting operators. \citet{koltchinskii2000} establish convergence of spectral
projections of the adjacency matrix to a limiting integral operator, with
similar results obtained using simplified proofs in
\citet{rosasco10}. \citet{vonluxburg2008} studies convergence of eigenvectors of
the Laplacian matrix for a neighborhood graph of fixed radius. \citet{belkin07} and
\citet{garciatrillos18} extend these results to the regime where the radius $r
\to 0$ as $n \to \infty$.

@@

At a high level, for PPR to be successful, the underlying density cluster must
be geometrically well-conditioned.  A basic requirement is that we need to avoid
clusters which contain arbitrarily thin bridges or spikes. As in the work of
\citet{chaudhuri2010}, we consider a thickened version of \smash{$\Cset \in
\Cbb_f(\lambda)$} defined as $\Csig := \set{x \in \Reals^d: \dist(x,\Cset) \leq
\sigma}$, which we call the \emph{$\sigma$-expansion} of $\Cset$. Here
\smash{$\dist(x,\Cset) := \inf_{y \in \Cset} \norm{y - x}$}.  We now list our
conditions on $\Csig$, which are defined with respect to $\sigma$ as well as positive numbers $\lambda_{\sigma}, \Lambda_{\sigma}, L$, $\theta \in (0,\infty)$, $\gamma \in [0,1]$, and $r \in (0,\frac{\sigma}{4d}]$.

\begin{enumerate}[label=(A\arabic*)]
\item
  \label{asmp: bounded_density}
  \emph{Bounded density within cluster:} The density is bounded above and below in $\mc{C}_{\sigma}$:
  $$
  \lambda_{\sigma} \leq \inf_{x \in \Csig} f(x) \leq \sup_{x \in \Csig} f(x)
  \leq \Lambda_{\sigma}.
  $$

\item 
  \label{asmp: low_noise_density}
  \emph{Low noise density:} For any $x \in \Rd$ such that $0 < \dist(x, \Csig) \leq \sigma$,     
  $$
  \inf_{x' \in \Csig} f(x') - f(x) \geq  \theta \cdot \dist(x, \Csig)^{\gamma}.  
  $$
  Roughly, this assumption ensures that the density decays sufficiently quickly
  as we move away from the target cluster $\Csig$, and is a standard assumption
  in the level-set estimation literature (see for instance \citet{singh2009}).

\item
  \label{asmp: embedding}
  \emph{Lipschitz embedding:}
  There exists some $g: \Reals^d \to \Reals^d$ satisfying each of the following:
  \begin{enumerate}
  \item the thickened set $\Csig = g(\mathcal{K})$, where $\mathcal{K}
    \subseteq \Rd$ is a convex set with diameter $\mathrm{diam}(\mathcal{K}) = \sup_{x,y \in \mathcal{K}}\norm{x - y} \leq \diam$;
  \item $\det(\nabla g (x)) = 1$ for all $x \in \mc{K}$, where $\nabla g(x)$ is
    the Jacobian of $g$ evaluated at $x;$ and 
  \item The restriction of $g$ to $\mc{K}$ is $L$-Lipschitz, i.e.
    $$
    \norm{g(x) - g(y)} \leq L \norm{x - y} ~
    \text{for all $x,y \in \mathcal{K}$}. 
    $$
  \end{enumerate}
  Succinctly, we assume that $\Csig$ is the image of a convex set with finite
  diameter under a measure preserving, Lipschitz transformation.

\item
  \label{asmp: bounded_volume}
  \emph{Bounded volume:}
  For a set $\mathcal{S} \subseteq \Reals^d$, define a $\Pbb$-weighted volume of 
  $\mathcal{S}$ to be  
  \begin{equation}
  \label{eqn:volume}
  \vol_{\Pbb,r}(\mathcal{S}) := \int_{\mathcal{S}} \Pbb(B(x,r)) f(x) \, dx.
  \end{equation}
  where $B(x,r)$ is the closed ball of radius $r$ centered at $x$. It holds that
  $$
  \vol_{\Pbb,r}(\Cset_\sigma) \leq \vol_{\Pbb,r}(\Reals^d \setminus \mc{C}_{\sigma}). 
  $$
\end{enumerate}

@@

\begin{definition}
  \label{def:well_conditioned_density_cluster}
  For $\lambda > 0$ and \smash{$\Cset \in \Cbb_f(\lambda)$}, let $\Cset$ satisfy  
  \ref{asmp: bounded_density}--\ref{asmp: bounded_volume} for some $\sigma, \lambda_{\sigma}, \Lambda_{\sigma}, L > 0$,  $\theta \in (0,\infty)$, $\gamma \in [0,1]$ and $r \in (0,\sigma/4d]$. Then, for positive universal constants $c_1$-$c_5$ to be specified later
  (in Theorems \ref{thm: conductance_upper_bound} and \ref{thm: mixing_time_upper_bound}), define
  \begin{align}
    \label{eqn: condition_number}
    \Phibf(\Cset) 
    & := c_1 r \frac{d}{\sigma} \frac{\lambda}{\lambda_{\sigma}}
    \frac{(\lambda_{\sigma} - \theta \frac{r^{\gamma}}{\gamma +
        1})}{\lambda_{\sigma}} \\
    \taubf(\Cset) & := c_2 \frac{\Lambda_{\sigma}^4 (d + 2)^3 
    	\rho^2 L^2}{\lambda_{\sigma}^4 r^2} \ln^2\Bigl(c_5^{1/d}\frac{\Lambda_{\sigma}^{2/d}L\rho}{2\lambda_{\sigma}^{2/d}r}\Bigr)
  \end{align}
  Letting \smash{$\kappa(\Cset) := \Phibf(\Cset) \cdot \taubf(\Cset)$},
  we call $\kappa(\Cset)$ the \emph{condition number} of $\Cset$.  We also call
  the set $\Cset$ a \emph{$\kappa$-well-conditioned} density cluster.  
\end{definition}

The condition number $\kappa(\Cset)$ succinctly captures the role of the various 
geometric parameters.\footnote{Technically speaking, the condition number depends on the numbers $\sigma, \lambda_{\sigma}, \Lambda_{\sigma},L > 0$, $\theta \in (0,\infty)$, $\gamma \in [0,1]$, and $r \in (0,\sigma/(4d))$ rather than on $\mc{C}$. However, we use the notation $\kappa(\mc{C})$ to make clear that the purpose of $\kappa$ is to measure the quality of the cluster $\mc{C}$.}  We note in passing that $\Phibf(\Cset)$ and
$\taubf(\Cset)$ are essentially the upper bounds  on \smash{$\Phi(\Csig[\Xbf];
  G_{n,r})$} and \smash{$\tau_{\infty}(G_{n,r}[\Csig[\Xbf]])$} that we derive in
our analysis later, in Section \ref{sec: analysis}.

@@

\begin{equation}
\begin{gathered}
\label{eqn: initialization}
0 < r \leq \frac{\sigma}{4d}, \quad 
\alpha \in {\textstyle [\frac{1}{3}, \frac{5}{12})} \cdot
\frac{1}{\taubf(\mc{C})}, \\ 
(L,U) \subseteq {\textstyle(\frac{1}{50},\frac{1}{5})} \cdot 
\frac{1}{2{n \choose 2} \vol_{\Pbb,r}(\Csig)}.  
\end{gathered}
\end{equation}

@@

We now present our first main result: a bound on the volume of the symmetric set
difference between the estimated cluster \smash{$\Cest$} and empirical cluster 
$\Csig[\Xbf]$. In this theorem, and hereafter, we let $c_i > 0$ denote
universal constants, and $b_i > 0$ denote constants which may depend on  
$\Pbb,\lambda,r,d$ and so on, but \emph{not} on the sample size 
$n$.

\begin{theorem}
  \label{thm: volume_ssd}
  Fix $\lambda,\sigma > 0$, let \smash{$\Cset \in \Cbb_f(\lambda)$} be a   
  $\kappa$-well-conditioned density cluster, and assume Algorithm \ref{alg: ppr}
  is well-initialized with respect to $\Cset$. Then for any 
  \begin{equation*}
  n \geq b_0 \cdot 
  \begin{cases*}
  \log(n)^{3/2},& ~~\textrm{if $d = 2$,} \\
  \log(n),& ~~\textrm{if $d \neq 2$}
  \end{cases*}
  \end{equation*}
  there exists a set $\Csig[\Xbf]^g 
  \subseteq \Csig[\Xbf]$ of large volume, \smash{$\vol_{n,r}(\Csig[\Xbf]^g)
    \geq \vol_{n,r}(\Csig[\Xbf])/2$}, such that the following holds: if 
  Algorithm~\ref{alg: ppr} is run with any seed node $v \in \Csig[\Xbf]^g$, then
  the PPR estimated cluster \smash{$\Cest$} satisfies
  \begin{equation}
    \label{eqn: volume_ssd}
    \Delta(\Cest, \Csig[\Xbf]) \leq c_0 \cdot \kappa(\Cset) \cdot
    \vol_{n,r}(\Csig[\Xbf]),  
  \end{equation}
  with probability at least $1 - b_0/n - c_0n\exp\{-b_{11}n\}$. 
\end{theorem}

@@

\begin{theorem}
  \label{thm: conductance_upper_bound}
  Fix $\lambda,\sigma>0$, and assume $\Cset \in \Cbb_f(\lambda)$ satisfies    
  Assumptions \ref{asmp: bounded_density}, \ref{asmp: low_noise_density}, and   
  \ref{asmp: bounded_volume}. Then for any $\delta \in (0,1/2)$, it holds that
  \begin{equation}
    \label{eqn: conductance_additive_error_bound}
    \frac{\Phi_{n,r}(\Csig[\Xbf])}{r} \leq (1 + 4\delta) c_1\frac{dr}{\sigma}
    \frac{\lambda}{\lambda_{\sigma}} \frac{(\lambda_{\sigma} -
      \theta\frac{r^{\gamma}}{\gamma+1})}{\lambda_{\sigma}},
  \end{equation}
  with probability at least $1 - 3\exp\set{-b_{10} n \delta^2}$. 
\end{theorem}

@@

\begin{theorem}
  \label{thm: mixing_time_upper_bound}
  Fix $\lambda, \sigma > 0$, and assume \smash{$\Cset \in \Cbb_f(\lambda)$} 
  satisfies Assumptions \ref{asmp: bounded_density} and \ref{asmp: embedding}.
  Also assume that \smash{$0 < r \leq \sigma/(4d)$}. Then for any \smash{$\delta \in \bigl(0,1/(8 + 16/d)\bigr)$}, and any
  \begin{equation}
  \label{eqn: mixing_time_sample_size}
  n \geq b_1 \cdot (1 + 4\delta) \cdot 
  \begin{cases*}
  \log(n)^{3/2},& ~~\textrm{if $d = 2$,} \\
  \log(n),& ~~\textrm{if $d \neq 2$}
  \end{cases*}
  \end{equation}
  the mixing time satisfies
  \begin{equation} 
    \label{eqn: mixing_time_upper_bound}
    \tau_{\infty}\bigl(G_{n,r}[\Csig[\Xbf]]\bigr) \leq (1 + c_3\delta) c_2 \frac{\Lambda_{\sigma}^4 (d + 2)^3 
      \rho^2 L^2}{\lambda_{\sigma}^4 r^2} \ln^2\Bigl(c_5^{1/d}\frac{\Lambda_{\sigma}^{2/d}L\rho}{2\lambda_{\sigma}^{2/d}r}\Bigr)
  \end{equation}
  with probability at least $1 - b_5/n - 4n\exp\{-\delta^2 b_2 n\}$.
\end{theorem}
The proof of Theorem \ref{thm: mixing_time_upper_bound} relies heavily on
analogous mixing time bounds developed for a continuous-space ``ball walk'' over
convex sets. To the best of our knowledge, our result is the first bound on the
mixing time of random walks over neighborhood graphs that is independent of $n$,
the number of vertices.

@@

\section{Proof of Theorem \ref{thm: conductance_upper_bound}}
\label{sec: proof_of_theorem_1}

@@

With this notation in place, the goal of Theorem~\ref{thm: conductance_upper_bound} is to show that for a universal constant $c_1 > 0$,
\begin{align*}
\Phi_{n,r}(\Csig[\Xbf]) := \frac{\cut_{n,r}(\Csig[\Xbf])}{\min\{\vol_{n,r}(\Csig[\Xbf]), \vol_{n,r}((\Rd \setminus \Csig)[\Xbf])\}} \leq c_1 \frac{d}{\sigma}
    \frac{\lambda}{\lambda_{\sigma}} \frac{(\lambda_{\sigma} -
      \theta\frac{r^{\gamma}}{\gamma+1})}{\lambda_{\sigma}}
\end{align*}
with probability at least $1 - 3\exp\{-nb\}$.

The proof of this theorem follows essentially from two technical Lemmas. 
Lemma~\ref{lem:ball_bounds_in_probability} relates the terms in the numerator and denominator of $\Phi_{n,r}(\Csig[\Xbf])$ to their expected values. We restate the conclusions of this Lemma: for any $\delta > 0$,
\begin{align}
\frac{\cut_{n,r}(\Csig[\Xbf])}{2{n \choose 2}} \leq (1 + \delta)\cut_{\Pbb,r}(\Csig) &, \quad \frac{\vol_{n,r}(\Csig[\Xbf])}{2{n \choose 2}} \geq (1 - \delta)\vol_{\Pbb,r}(\Csig) \nonumber \\ \frac{\vol_{n,r}((\Rd \setminus \Csig)[\Xbf])}{2{n \choose 2}}& \geq (1 - \delta)\vol_{\Pbb,r}(\Rd \setminus \Csig) \label{eqn: conductance_upper_bound_pf1}
\end{align}
with probability at least 
\begin{align*}
1 - & \exp\set{-n \delta^2 (\cut_{\Pbb,r}(\Csig))^2} - \exp\set{-n \delta^2 (\vol_{\Pbb,r}(\Csig))^2} - \exp\set{-n \delta^2 (\vol_{\Pbb,r}(\Rd\setminus \Csig))^2} \\
& \geq 1 - 3\exp\set{-n \delta^2 (\cut_{\Pbb,r}(\Csig))^2}.
\end{align*} 
We note that as a consequence of~\ref{asmp: bounded_volume} we have that $\vol_{\Pbb,r}(\Rd \setminus \Csig)  \geq \vol_{\Pbb,r}(\Csig)$, so it will suffice to lower bound $\vol_{\Pbb,r}(\Csig)$ (since a lower bound for $\vol_{\Pbb,r}(\Rd \setminus \Csig)$ follows).
The following result provides upper and lower bounds on the expected values $\cut_{\Pbb,r}(\Csig)$ and $\vol_{\Pbb,r}(\Csig)$ respectively:
\begin{lemma}
	\label{lem: expected_density_cut}\label{lem: expected_density_volume}
	Under the setup and conditions of Theorem \ref{thm: conductance_upper_bound}, and for any $0 < r \leq \sigma/(2d)$,
	\begin{align}
	\cut_{\Pbb,r}(\Csig) &\leq \frac{ d \nu_d r^{d+1} \lambda}{\sigma - dr} \left(\lambda_{\sigma} - \theta\frac{r^{\gamma}}{\gamma + 1}\right) \nu(\Csig), \label{eqn:claim_one} \\
	\vol_{\Pbb,r}(\Csig) &\geq \biggl(1 + \frac{dr}{\sigma - dr}\biggr)^{-1} \lambda_{\sigma}^2 \nu_d r^d \nu(\Csig).\label{eqn:claim_two}
	\end{align}
\end{lemma}
Taking Lemma~\ref{lem: expected_density_cut} and \eqref{eqn: conductance_upper_bound_pf1} as given we can now complete the proof of the theorem. We lower bound $\Phi_{n,r}(\Csig[\Xbf])$ as follows:
\begin{equation*}
\Phi_{n,r}(\Csig[\Xbf]) \leq \frac{(1 + \delta)\cut_{\Pbb,r}(\Csig)}{(1 - \delta)\vol_{\Pbb,r}(\Csig)} \leq \frac{(1 + \delta)}{(1 - \delta)(1 - dr/\sigma)^2}\frac{d r \lambda \left(\lambda_{\sigma} - \theta\frac{r^{\gamma}}{\gamma + 1}\right)}{\sigma \lambda_{\sigma}^2}.
\end{equation*}
Noting that $(1 - dr/\sigma)^{-2} \leq 16/9$ for all $r \leq \sigma/(4d)$, the theorem is satisfied by choosing $c_1 = 16/9$ and $b = (\cut_{\Pbb,r}(\Csig))^2$.
